---
title: "Tooth Growth Data Analysis"
author: "Christopher Hair"
date: "September 15, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(huxtable)
```

# Overview

In this article we will explore the `ToothGrowth` data that is included in the `datasets` package in R. This dataset is used to show the effects of Vitamin C on Tooth Growth in Guinea Pigs. We will show some information about what the data looks like and how the data is structured. We can then look at some basic analysis on comparing the type of supplement and dosage of each supplement. Using this analysis we will attempt to draw a hypothesis to compare which component supplement and dosage has the larger impact on the tooth growth.

# Summarizing the Data

First we need to load the `datasets` library and then we can look at some of the basic properties of the `ToothGrowth` sample data and how the information is structured as well as some introduction into the types of values we are are going to evaluate in our testing.

```{r load_data}
library(datasets)

str(ToothGrowth)
head(ToothGrowth)
summary(ToothGrowth)
```

# Basic Analysis

In this section we will go through our two major points of 

## 

```{r supp_summary}
grouped_summay <- group_by(ToothGrowth, supp, dose) %>%
    summarise(group_count=n(),
              group_mean=mean(len),
              group_med=median(len),
              group_sd=sd(len))

as_hux(grouped_summay %>%
           select("Supplement"=supp,
                  "Dosage"=dose,
                  "Count"=group_count,
                  "Mean"=group_mean,
                  "Median"=group_med,
                  "Standard Deviation"=group_sd), 
       add_colnames=TRUE) %>%
    set_bold(1, everywhere, TRUE) %>%
    set_all_borders(1)
```

What we can see here is that orange juice provides a slightly larger average impact on tooth growth compared to vitamin C alone. If we visualize the data we can see similar results.

```{r data_plot}
ggplot(data = ToothGrowth) +
    geom_boxplot(aes(fill = factor(dose), x = factor(dose), y = len)) +
    facet_grid(.~supp) +
    ggtitle("Tooth Growth by Supplement and Dosage") +
    xlab("Dosage") +
    ylab("Tooth Growth") +
    theme(legend.position="none",
          plot.title = element_text(hjust = 0.5))

ggplot(data = ToothGrowth) +
    geom_boxplot(aes(fill = supp, x = supp, y = len)) +
    facet_grid(.~dose) +
    ggtitle("Tooth Growth by Dosage and Supplement") +
    xlab("Supplement") +
    ylab("Tooth Growth") +
    theme(legend.position = "none",
          plot.title = element_text(hjust = 0.5))
```

# Hypothesis Testing

From the basic analysis we can now see that we now have the following hypothesis statement to work with.

- **H~0~** Orange Juice has a greater impact on tooth growth in guinea pigs than Vitamin C
- **H~a~** Orange Juise has a lesser impact on tooth growth in guinea pigs than Viatmin C

